---
title: "p8105_hw2_xc2472"
author: "Xing Chen"
date: 2019-10-1
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```


# Problem 1

## Mr. Trash Wheel dataset

```{r}
trashwheel_data = 
  read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
             sheet = "Mr. Trash Wheel", 
             range = "A2:N408")

trashwheel_data = janitor::clean_names(trashwheel_data)

trashwheel_data = drop_na(trashwheel_data, dumpster)

trashwheel_data = mutate(trashwheel_data, 
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)
)
```

## The precipitation dataset

import the datasets

```{r}
prcp17_data = 
  read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2017 Precipitation",
             range = "A2:B14") %>% 
  janitor::clean_names()

prcp18_data = 
  read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2018 Precipitation",
             range = "A2:B14") %>% 
  janitor::clean_names()
```

omit rows without precipitation data and add a variable year

```{r}
prcp17_data = drop_na(prcp17_data, total) %>% 
  mutate(year = 2017)

prcp18_data = drop_na(prcp18_data, total) %>% 
  mutate(year = 2018)
```

Combine the precipitation datasets and convert "month" variable to character variable

```{r}
prcp_combine_data = 
  bind_rows(prcp17_data, prcp18_data) %>% 
  mutate(
    month = month.name[month])
```

**Description**


# Problem 2

## read and clean the pols-month.csv

the frile plos_month is read and cleaned. 

the variable `mon` is separated into 3 variables, and the value in `month` varriable is changed to character.

A president variable is created and drop the "day" variable

```{r}
pols_month = 
  read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    month = as.numeric(month),
    month = month.name[month]) %>% 
  pivot_longer(
    cols = starts_with("prez"),
    names_to = "president",
    names_prefix = "prez_",
    values_to = "time") %>% 
  filter(time != 0) %>% 
  select(-day)
```


## read and clean the snp.csv

the file snp is read and cleaned.

the variable `date` is separated into 3 variables `year`, `month`, and `day`
then, the variable `day` is dropped and the dataset is arranged so that `year` and
`month` is leading the column, for consistency across datasets. 

```{r}
snp = 
  read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(
    month = as.numeric(month),
    month = month.name[month]) %>% 
  select(year, month, -day, close)
```


## read and clean the unemployment.csv

```{r}
unemploy = 
  read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemploy_rate")
```






