---
title: "p8105_hw2_xc2472"
author: "Xing Chen"
date: 2019-10-1
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```


# Problem 1

## Mr. Trash Wheel dataset

```{r}
trashwheel_data = 
  read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
             sheet = "Mr. Trash Wheel", 
             range = "A2:N408")

trashwheel_data = janitor::clean_names(trashwheel_data)

trashwheel_data = drop_na(trashwheel_data, dumpster)

trashwheel_data = mutate(trashwheel_data, 
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)
)
```

## The precipitation dataset

import the datasets

```{r}
prcp17 = 
  read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2017 Precipitation",
             range = "A2:B14")
prcp17 = janitor::clean_names(prcp17)

prcp18 = 
  read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2018 Precipitation",
             range = "A2:B14")
prcp18 = janitor::clean_names(prcp18)
```

omit rows without precipitation data and add a variable year

```{r}
prcp17 = drop_na(prcp17, total)

prcp17 = mutate(prcp17, year = 2017)

prcp18 = drop_na(prcp18, total)

prcp18 = mutate(prcp18, year = 2018)
```

Combine the precipitation datasets and convert "month" variable to character variable

```{r}
prcp_combine = bind_rows(prcp17, prcp18)
# haven't change the month variable

```

**Description**


# Problem 2

## read and clean the pols-month.csv

```{r}
pols_month = 
  read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-")
# haven't change the month to its names yet
```

create a president variable and drop the "day" variable

```{r}
#this is not right
pols_month = pivot_longer(pols_month, 
                          cols = starts_with("prez"), 
                          names_to = "president",
                          names_prefix = "prez_",
                          )
pols_month = select(pols_month, -day)
```


## read and clean the snp.csv

```{r}
snp = 
  read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  select(year, month, -day, close)
```


## read and clean the unemployment.csv

```{r}
unemploy = 
  read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemploy_rate")
```






